---
title: "R Notebook"
output: html_notebook
---

Phil Ramontal
CIS 627
NFL Analysis 




```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(corrplot)
library(caTools)
library(caret)
library(ROCR)
library(MASS)
library(randomForest)
library(neuralnet)
library(rpart)
library(e1071)
library(tree)
```


```{r}
#Import dataset
nfl_data<-read_csv("NFL Play by Play 2009-2018.csv")
```


```{r}
#omit all NA's from dataset
nfl_na<-drop_na(nfl_data)
```
  

```{r}
nfl1<-subset(nfl_na, play_type == "pass")
nfl2<-subset(nfl_na,  play_type == "run")
nflDataa<-rbind(nfl1,  nfl2)

print(nflDataa)
```

```{r}
numeric.var <- sapply(nflDataa, is.numeric)

corrplot(cor(nflDataa[,numeric.var]), method = "number", number.cex = 0.7)
```


```{r}
nflDataa$yards_gained<-as.integer(nflDataa$yards_gained)
nflDataa$down<-as.integer(nflDataa$down)
nflDataa$ydsnet<-as.integer(nflDataa$ydsnet)
nflDataa$yardline_100<-as.integer(nflDataa$yardline_100)
```


```{r}
#What playtype should be excuted between each 10 yard increment on down 1-4
yard100<-subset(nflDataa, yardline_100 >=50 & down == "1")
yard50<-subset(nflDataa, yardline_100 >=0 & yardline_100 <50 & down == "1")

```
```{r}
yard100_2<-subset(nflDataa, yardline_100 >=50 & down == "2")
yard50_2<-subset(nflDataa, yardline_100 >=0 & yardline_100 <50 & down == "2")

```
```{r}
yard100_3<-subset(nflDataa, yardline_100 >=50 & down == "3")
yard50_3<-subset(nflDataa, yardline_100 >=0 & yardline_100 <50 & down == "3")

```
```{r}
yard100_4<-subset(nflDataa, yardline_100 >=50 & down == "4")
yard50_4<-subset(nflDataa, yardline_100 >=0 & yardline_100 <50 & down == "4")

```

# Yard 100-50 Best Play Type prediction Down 1
```{r}
set.seed(123)
yard100<-subset(nflDataa, yardline_100 >=50 & down == "1")
yard100$play_type<-as.factor(yard100$play_type)
sample<-sample.split(yard100$play_type, SplitRatio = .70)
training<-subset(yard100, sample == TRUE)
print(training)
testing<-subset(yard100, sample == FALSE)
print(testing)
```


```{r}
#Support Vector Machine Model down 1 
svm_nfl <- svm(play_type ~ yards_gained + yardline_100 + down + qtr + ydsnet, data = training, prob.model = TRUE)


plot(svm_nfl, training, ydstogo~yards_gained)

svm_pred<- predict(svm_nfl,newdata = testing)

tb<- table(testing$play_type, svm_pred)
tb
results_svm<-data.frame(Actual = testing$play_type, Prediction = svm_pred)
confusionMatrix(tb)

tn<-subset(results_svm, Actual == "run" & Prediction == "run")
tp<-subset(results_svm, Actual == "pass" & Prediction == "pass")
fp<-subset(results_svm, Actual == "pass" & Prediction == "run")
fn<-subset(results_svm, Actual == "run"& Prediction == "pass")
svm_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
svm_tp<-ifelse(results_svm$Actual == results_svm$Prediction, 1,0)
svm_tp<-as.data.frame(svm_tp)
svm_tp$model<-"svm"
names(svm_tp)[1] = "accuracy"
```

```{r}
#ANN model down 1
nfl_nn<-neuralnet(play_type ~ yards_gained + yardline_100 +  down + qtr + ydsnet ,data = training, hidden =2)

```


```{r}
nfl_nn$result.matrix
plot(nfl_nn)
#Prediction

temp_test<-subset(testing,select = c("yards_gained", "yardline_100", "down", "ydsnet", "qtr"))
head(temp_test)       
nfl_nn.results<-compute(nfl_nn,temp_test)
nfl_nn.results$net.result
prob<-nfl_nn.results$net.result
#nn_pred<-predict(nfl_nn, newdata =  testing, type = "matrix" )
ann<-ifelse(prob>=.5, "run", "pass")
ann<-as.vector(ann)
resultsann<-data.frame(Actual = testing$play_type, Prediction = ann)
tb_ann<-table(resultsann$Actual, resultsann$Prediction)


confusionMatrix(tb_ann)

tn<-subset(resultsann, Actual == "run" & Prediction == "run")
ann_tp<-subset(resultsann, Actual == "pass" & Prediction == "pass")
fp<-subset(resultsann, Actual == "pass" & Prediction == "run")
fn<-subset(resultsann, Actual == "run"& Prediction == "pass")
ann_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
ann_tp<-ifelse(resultsann$Actual == resultsann$Prediction, 1,0)
ann_tp<-as.data.frame(ann_tp)
ann_tp$model<-"ann"
names(ann_tp)[1] = "accuracy"
```


```{r}
#RandonForest model down 1
library(party)
random_nfl<-randomForest(play_type ~ yards_gained + yardline_100 + qtr + ydsnet ,data = training)
random_nfl

plot(random_nfl)


#Prediction
ranForest_pred<-predict(random_nfl, testing)

ran_tb<-table(testing$play_type, ranForest_pred, dnn = c("Actual", "Prediction"))
results_rf<-data.frame(Actual = testing$play_type, Prediction = ranForest_pred)

confusionMatrix(ran_tb)

tn<-subset(results_rf, Actual == "run"& Prediction == "run")
rf_tp<-subset(results_rf, Actual == "pass"&Prediction == "pass")
fp<-subset(results_rf, Actual == "pass"& Prediction == "run")
fn<-subset(results_rf, Actual == "run"& Prediction == "pass")
rf_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
rf_tp<-ifelse(results_rf$Actual == results_rf$Prediction, 1,0)
rf_tp<-as.data.frame(rf_tp)
rf_tp$model<-"rf"
names(rf_tp)[1] = "accuracy"
```


```{r}
compare<-data.frame(method = c("Support Vector Machine", "Neural Network", "Random Forest"), accuracy = c(svm_accuracy, ann_accuracy, rf_accuracy))
ggplot(data = compare, mapping = aes(x= method, y = accuracy, fill = method)) +geom_bar(stat = "identity")+geom_text(aes(label = round(accuracy,4)), position = position_stack(vjust = .5), color = "white") + labs (x = "Method", y = "Accuracy", title = "Accuracy for Down 1 Yard Line 100-50")
```

```{r}
truepositive<-rbind(svm_tp,ann_tp,rf_tp)
truepositive<-as.data.frame(truepositive)
truepositive$model<-as.factor(truepositive$model)
anova_model<-aov(accuracy~model, data = truepositive)
summary(anova_model)
```



# Yard Line 50-0 Best Play Type prediction Down 1
```{r}
set.seed(123)
yard50<-subset(nflDataa, yardline_100 >=0 & yardline_100 <50 & down == "1")
yard50$play_type<-as.factor(yard50$play_type)
sample<-sample.split(yard50$play_type, SplitRatio = .70)
training1<-subset(yard50, sample == TRUE)
testing1<-subset(yard50, sample == FALSE)
```

```{r}
#Support Vector Machine Model #2 Down 1 Yard 0-50
svm_nfl1 <- svm(play_type ~ yards_gained + yardline_100 + down  + qtr + ydsnet, data = training1, prob.model = TRUE)
svm_nfl1

plot(svm_nfl1, training1, ydstogo~yards_gained)

svm_pred1<- predict(svm_nfl1,newdata = testing1)

tab<- table(testing1$play_type, svm_pred1)

results_svm1<-data.frame(Actual = testing1$play_type, Prediction = svm_pred1)

confusionMatrix(tab)

tn<-subset(results_svm1, Actual == "run"& Prediction == "run")
tp<-subset(results_svm1, Actual == "pass"& Prediction == "pass")
fp<-subset(results_svm1, Actual == "pass"& Prediction == "run")
fn<-subset(results_svm1, Actual == "run"& Prediction == "pass")
svm1_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
svm_tp1<-ifelse(results_svm1$Actual == results_svm1$Prediction, 1,0)
svm_tp1<-as.data.frame(svm_tp1)
svm_tp1$model1<-"SVM1"
names(svm_tp1)[1] = "accuracy1"
```

```{r}
#ANN model Down 1 Yard 0-50
nfl_nn1<-neuralnet(play_type ~ yards_gained + yardline_100 +  down + qtr + ydsnet ,data = training1, hidden =2)


nfl_nn1$result.matrix
 plot(nfl_nn1)
#Prediction

temp_test1<-subset(testing1,select = c("yards_gained", "yardline_100", "down", "ydsnet", 'qtr'))
head(temp_test1)       
nfl_nn.results1<-compute(nfl_nn1,temp_test1)
nfl_nn.results1$net.result
prob1<-nfl_nn.results1$net.result
ann1<-ifelse(prob1>=.5, "run", "pass")
ann1<-as.vector(ann1)
resultsann1<-data.frame(Actual = testing1$play_type, Prediction = ann1)
tb_ann1<-table(resultsann1$Actual, resultsann1$Prediction)

confusionMatrix(tb_ann)

tn<-subset(resultsann1, Actual == "run"& Prediction == "run")
tp<-subset(resultsann1, Actual == "pass"& Prediction == "pass")
fp<-subset(resultsann1, Actual == "pass"& Prediction == "run")
fn<-subset(resultsann1, Actual == "run"& Prediction == "pass")
ann1_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
ann_tp1<-ifelse(resultsann1$Actual == resultsann1$Prediction, 1,0)
ann_tp1<-as.data.frame(ann_tp1)
ann_tp1$model1<-"ANN1"
names(ann_tp1)[1] = "accuracy1"
```

```{r}
random_nfl1<-randomForest(play_type ~ yards_gained + yardline_100 + down + qtr + ydsnet, method = "anova",  data = training1)
random_nfl1

plot(random_nfl1)

#Prediction
ranForest_pred1<-predict(random_nfl1, testing1)

ran_tb1<-table(testing1$play_type, ranForest_pred1, dnn = c("Actual", "Prediction"))
results_rf1<-data.frame(Actual = testing1$play_type, Prediction = ranForest_pred1)


confusionMatrix(ran_tb1)

tn<-subset(results_svm, Actual == "run"& Prediction == "run")
tp<-subset(results_svm, Actual == "pass"& Prediction == "pass")
fp<-subset(results_svm, Actual == "pass"& Prediction == "run")
fn<-subset(results_svm, Actual == "run"& Prediction == "pass")
rf1_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
rf_tp1<-ifelse(results_rf1$Actual == results_rf1$Prediction, 1,0)
rf_tp1<-as.data.frame(rf_tp1)
rf_tp1$model1<-"RF1"
names(rf_tp1)[1] = "accuracy1"
```

```{r}
compare1<-data.frame(method = c("Support Vector Machine", "Neural Network", "Random Forest"), accuracy = c(svm1_accuracy, ann1_accuracy, rf1_accuracy))
ggplot(data = compare1, mapping = aes(x= method, y = accuracy, fill = method)) +geom_bar(stat = "identity")+geom_text(aes(label = round(accuracy,4)), position = position_stack(vjust = .5), color = "white") + labs (x = "Method", y = "Accuracy", title = "Accuracy for Down 1 Yard Line 50-0")
```


# Yard 100-50 Best Play Type prediction Down 2
```{r}
set.seed(123)
yard100_2$play_type<-as.factor(yard100_2$play_type)
sample<-sample.split(yard100_2$play_type, SplitRatio = .70)
training2<-subset(yard100_2, sample == TRUE)
testing2<-subset(yard100_2, sample == FALSE)
```

```{r}
#Support Vector Machine Model 1 down 2 Yard 100-50
svm_nfl2 <- svm(play_type ~ yards_gained + yardline_100 + down + qtr + ydsnet, data = training2, prob.model = TRUE)
svm_nfl2

plot(svm_nfl2, training2, ydstogo~yards_gained)

svm_pred2<- predict(svm_nfl2,newdata = testing2)

tab2<- table(Actual = testing2$play_type, svm_pred2)

results_svm2<-data.frame(Actual = testing2$play_type, Prediction = svm_pred2)

confusionMatrix(tab2)

tn<-subset(results_svm2, Actual == "run"& Prediction == "run")
tp<-subset(results_svm2, Actual == "pass"& Prediction == "pass")
fp<-subset(results_svm2, Actual == "pass"& Prediction == "run")
fn<-subset(results_svm2, Actual == "run"& Prediction == "pass")
svm2_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
svm_tp2<-ifelse(results_svm2$Actual == results_svm2$Prediction, 1,0)
svm_tp2<-as.data.frame(svm_tp2)
svm_tp1$model<-"SVM2"
names(svm_tp2)[1] = "accuracy"
```

```{r}
#ANN model down 2 Yard 100-50
nfl_nn2<-neuralnet(play_type ~ yards_gained + yardline_100 +  down + qtr + ydsnet ,data = training2, hidden =2)


nfl_nn2$result.matrix
plot(nfl_nn2)
#Prediction

temp_test2<-subset(testing2,select = c("yards_gained", "yardline_100", "down", "qtr", "ydsnet"))
head(temp_test2)       
nfl_nn.results2<-compute(nfl_nn2,temp_test2)
nfl_nn.results2$net.result
prob2<-nfl_nn.results2$net.result
ann2<-ifelse(prob2>=.5, "run", "pass")
ann2<-as.vector(ann2)
resultsann2<-data.frame(Actual = testing2$play_type, Prediction = ann2)
tb_ann2<-table(resultsann2$Actual, resultsann2$Prediction)


confusionMatrix(tb_ann2)

tn<-subset(resultsann2, Actual == "run"& Prediction == "run")
tp<-subset(resultsann2, Actual == "pass"& Prediction == "pass")
fp<-subset(resultsann2, Actual == "pass"& Prediction == "run")
fn<-subset(resultsann2, Actual == "run"& Prediction == "pass")
ann2_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
ann_tp2<-ifelse(resultsann2$Actual == resultsann2$Prediction, 1,0)
ann_tp2<-as.data.frame(ann_tp2)
ann_tp2$model1<-"ANN2"
names(ann_tp2)[1] = "accuracy"
```

```{r}
#RandomForest model Down 2 yard 100-50
random_nfl2<-randomForest(play_type ~ yards_gained + yardline_100 + down + qtr + ydsnet  ,data = training2)
random_nfl2

plot(random_nfl2)

#Prediction
ranForest_pred2<-predict(random_nfl2, testing2)

ran_tb2<-table(testing2$play_type, ranForest_pred2, dnn = c("Actual", "Prediction"))
results_rf2<-data.frame(Actual = testing2$play_type, Prediction = ranForest_pred2)


confusionMatrix(ran_tb2)

tn<-subset(results_rf2, Actual == "run"& Prediction == "run")
tp<-subset(results_rf2, Actual == "pass"& Prediction == "pass")
fp<-subset(results_rf2, Actual == "pass"& Prediction == "run")
fn<-subset(results_rf2, Actual == "run"& Prediction == "pass")
rf2_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
rf_tp2<-ifelse(results_rf1$Actual == results_rf1$Prediction, 1,0)
rf_tp2<-as.data.frame(rf_tp1)
rf_tp2$model<-"RF2"
names(rf_tp2)[1] = "accuracy"
```

```{r}
#Comparing Accuracies
compare2<-data.frame(method = c("Support Vector Machine", "Neural Network", "Random Forest"), accuracy = c(svm2_accuracy, ann2_accuracy, rf2_accuracy))
ggplot(data = compare2, mapping = aes(x= method, y = accuracy, fill = method)) +geom_bar(stat = "identity")+geom_text(aes(label = round(accuracy,4)), position = position_stack(vjust = .5), color = "white") + labs (x = "Method", y = "Accuracy", title = "Accuracy for Down 2 Yard Line 100-50")
```


# Yard 0-50 Best Play Type prediction Down 2
```{r}
set.seed(123)
yard50_2$play_type<-as.factor(yard50_2$play_type)
sample<-sample.split(yard50_2$play_type, SplitRatio = .70)
training3<-subset(yard50_2, sample == TRUE)
testing3<-subset(yard50_2, sample == FALSE)
```

```{r}
#Support Vector Machine Model Down 2 Yard 0-50
svm_nfl3 <- svm(play_type ~ yards_gained + yardline_100 + down + qtr + ydsnet, data = training3, prob.model = TRUE)
svm_nfl3

plot(svm_nfl3, training3, ydstogo~yards_gained)

svm_pred3<- predict(svm_nfl3,newdata = testing3)

tab3<- table(testing3$play_type, svm_pred3)

results_svm3<-data.frame(Actual = testing3$play_type, Prediction = svm_pred3)

confusionMatrix(tab3)

tn<-subset(results_svm3, Actual == "run"& Prediction == "run")
tp<-subset(results_svm3, Actual == "pass"& Prediction == "pass")
fp<-subset(results_svm3, Actual == "pass"& Prediction == "run")
fn<-subset(results_svm3, Actual == "run"& Prediction == "pass")
svm3_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
svm_tp3<-ifelse(results_svm3$Actual == results_svm3$Prediction, 1,0)
svm_tp3<-as.data.frame(svm_tp3)
svm_tp3$model<-"SVM3"
names(svm_tp3)[1] = "accuracy"
```

```{r}
#ANN model Down 2 Yard 0-50
nfl_nn3<-neuralnet(play_type ~ yards_gained + yardline_100 +  down + qtr + ydsnet ,data = training3, hidden =2)

  
nfl_nn3$result.matrix
plot(nfl_nn3)
#Prediction

temp_test3<-subset(testing3,select = c("yards_gained", "yardline_100", "down", "qtr", "ydsnet"))
head(temp_test3)       
nfl_nn.results3<-compute(nfl_nn3,temp_test3)
nfl_nn.results3$net.result
prob3<-nfl_nn.results3$net.result
ann3<-ifelse(prob3>=.5, "run", "pass")
ann3<-as.vector(ann3)
resultsann3<-data.frame(Actual = testing3$play_type, Prediction = ann3)
tb_ann3<-table(resultsann3$Actual, resultsann3$Prediction)


confusionMatrix(tb_ann3)

tn<-subset(resultsann3, Actual == "run"& Prediction == "run")
tp<-subset(resultsann3, Actual == "pass"& Prediction == "pass")
fp<-subset(resultsann3, Actual == "pass"& Prediction == "run")
fn<-subset(resultsann3, Actual == "run"& Prediction == "pass")
ann3_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
ann_tp3<-ifelse(resultsann3$Actual == resultsann3$Prediction, 1,0)
ann_tp3<-as.data.frame(ann_tp3)
ann_tp3$model<-"ANN3"
names(ann_tp3)[1] = "accuracy"
```

```{r}
random_nfl3<-randomForest(play_type ~ yards_gained + yardline_100 + down  + qtr + ydsnet ,data = training3)
random_nfl3

plot(random_nfl3)

#Prediction
ranForest_pred3<-predict(random_nfl3, testing3)

ran_tb3<-table(testing3$play_type, ranForest_pred3, dnn = c("Actual", "Prediction"))
results_rf3<-data.frame(Actual = testing3$play_type, Prediction = ranForest_pred3)


confusionMatrix(ran_tb3)

tn<-subset(results_svm, Actual == "run"& Prediction == "run")
tp<-subset(results_svm, Actual == "pass"& Prediction == "pass")
fp<-subset(results_svm, Actual == "pass"& Prediction == "run")
fn<-subset(results_svm, Actual == "run"& Prediction == "pass")
rf3_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
rf_tp3<-ifelse(results_rf3$Actual == results_rf3$Prediction, 1,0)
rf_tp3<-as.data.frame(rf_tp3)
rf_tp3$model<-"RF3"
names(rf_tp3)[1] = "accuracy"
```

```{r}
#Comparing Accuracies
compare3<-data.frame(method = c("Support Vector Machine", "Neural Network", "Random Forest"), accuracy = c(svm3_accuracy, ann3_accuracy, rf3_accuracy))
ggplot(data = compare3, mapping = aes(x= method, y = accuracy, fill = method)) +geom_bar(stat = "identity")+geom_text(aes(label = round(accuracy,4)), position = position_stack(vjust = .5), color = "white") + labs (x = "Method", y = "Accuracy", title = "Accuracy for Down 2 Yard Line 50-0")
```


# Yard 100-50 Best Play Type prediction Down 3

```{r}
set.seed(123)
yard100_3$play_type<-as.factor(yard100_3$play_type)
sample<-sample.split(yard100_3$play_type, SplitRatio = .70)
training4<-subset(yard100_3, sample == TRUE)
testing4<-subset(yard100_3, sample == FALSE)
```

```{r}
#Support Vector Machine Model down 3 yard 100-50
svm_nfl4 <- svm(play_type ~ yards_gained + yardline_100 + down + qtr + ydsnet, data = training4, prob.model = TRUE)
svm_nfl4

plot(svm_nfl4, training4, ydstogo~yards_gained)

svm_pred4<- predict(svm_nfl4,newdata = testing4)

tab4<- table(testing4$play_type, svm_pred4)

results_svm4<-data.frame(Actual = testing4$play_type, Prediction = svm_pred4)

confusionMatrix(tab4)

tn<-subset(results_svm4, Actual == "run", Prediction == "run")
tp<-subset(results_svm4, Actual == "pass", Prediction == "pass")
fp<-subset(results_svm4, Actual == "pass", Prediction == "run")
fn<-subset(results_svm4, Actual == "run", Prediction == "pass")
svm4_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
svm_tp4<-ifelse(results_svm4$Actual == results_svm4$Prediction, 1,0)
svm_tp4<-as.data.frame(svm_tp4)
svm_tp4$model<-"SVM4"
names(svm_tp4)[1] = "accuracy"
```

```{r}
nfl_nn4<-neuralnet(play_type ~ yards_gained + yardline_100 +  down + qtr + ydsnet ,data = training4, hidden =2)


nfl_nn4$result.matrix
plot(nfl_nn4)
#Prediction

temp_test4<-subset(testing4,select = c("yards_gained", "yardline_100", "down", "qtr", "ydsnet"))
head(temp_test4)       
nfl_nn.results4<-compute(nfl_nn4,temp_test4)
nfl_nn.results4$net.result
prob4<-nfl_nn.results4$net.result
ann4<-ifelse(prob4>=.5, "run", "pass")
ann4<-as.vector(ann4)
resultsann4<-data.frame(Actual = testing4$play_type, Prediction = ann4)
tb_ann4<-table(resultsann4$Actual, resultsann4$Prediction)


confusionMatrix(tb_ann4)

tn<-subset(resultsann4, Actual == "run", Prediction == "run")
tp<-subset(resultsann4, Actual == "pass", Prediction == "pass")
fp<-subset(resultsann4, Actual == "pass", Prediction == "run")
fn<-subset(resultsann4, Actual == "run", Prediction == "pass")
ann4_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
ann_tp4<-ifelse(resultsann4$Actual == resultsann4$Prediction, 1,0)
ann_tp4<-as.data.frame(ann_tp4)
ann_tp4$model<-"ANN4"
names(ann_tp4)[1] = "accuracy"
```

```{r}
#RandonForest model down 3 Yard 100-50
random_nfl4<-randomForest(play_type ~ yards_gained + yardline_100 + down + qtr + ydsnet  ,data = training4)
random_nfl4

plot(random_nfl4)

#Prediction
ranForest_pred4<-predict(random_nfl4, testing4)

ran_tb4<-table(testing4$play_type, ranForest_pred4, dnn = c("Actual", "Prediction"))
results_rf4<-data.frame(Actual = testing4$play_type, Prediction = ranForest_pred4)


confusionMatrix(ran_tb4)

tn<-subset(results_svm, Actual == "run", Prediction == "run")
tp<-subset(results_svm, Actual == "pass", Prediction == "pass")
fp<-subset(results_svm, Actual == "pass", Prediction == "run")
fn<-subset(results_svm, Actual == "run", Prediction == "pass")
rf4_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
rf_tp4<-ifelse(results_rf4$Actual == results_rf4$Prediction, 1,0)
rf_tp4<-as.data.frame(rf_tp4)
rf_tp4$model<-"RF4"
names(rf_tp4)[1] = "accuracy"
```
```{r}
#Comparing Accuracies
compare4<-data.frame(method = c("Support Vector Machine", "Neural Network", "Random Forest"), accuracy = c(svm4_accuracy, ann4_accuracy, rf4_accuracy))
ggplot(data = compare4, mapping = aes(x= method, y = accuracy, fill = method)) +geom_bar(stat = "identity")+geom_text(aes(label = round(accuracy,4)), position = position_stack(vjust = .5), color = "white") + labs (x = "Method", y = "Accuracy", title = "Accuracy for Down 2 Yard Line 100-50")
```

# Yard 0-50 Best Play Type prediction Down 3

```{r}
set.seed(123)
yard50_3$play_type<-as.factor(yard50_3$play_type)
sample<-sample.split(yard50_3$play_type, SplitRatio = .70)
training5<-subset(yard50_3, sample == TRUE)
testing5<-subset(yard50_3, sample == FALSE)
```


```{r}
#Support Vector Machine Model 1 down 3 Yard 0-50
svm_nfl5 <- svm(play_type ~ yards_gained + yardline_100 + down+ qtr + ydsnet  , data = training5, prob.model = TRUE)
svm_nfl5

plot(svm_nfl5, training4, ydstogo~yards_gained)

svm_pred5<- predict(svm_nfl5,newdata = testing5)

tab5<- table(testing5$play_type, svm_pred5)

results_svm5<-data.frame(Actual = testing5$play_type, Prediction = svm_pred5)

confusionMatrix(tab5)

tn<-subset(results_svm, Actual == "run", Prediction == "run")
tp<-subset(results_svm, Actual == "pass", Prediction == "pass")
fp<-subset(results_svm, Actual == "pass", Prediction == "run")
fn<-subset(results_svm, Actual == "run", Prediction == "pass")
svm5_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
svm_tp5<-ifelse(results_svm5$Actual == results_svm5$Prediction, 1,0)
svm_tp5<-as.data.frame(svm_tp5)
svm_tp5$model<-"SVM5"
names(svm_tp5)[1] = "accuracy"
```

```{r}
#ANN model Down 3 Yard 0-50
nfl_nn5<-neuralnet(play_type ~ yards_gained + yardline_100 +  down + qtr + ydsnet ,data = training5, hidden =2)


nfl_nn5$result.matrix
plot(nfl_nn5)
#Prediction

temp_test5<-subset(testing5,select = c("yards_gained", "yardline_100", "down", "qtr", "ydsnet"))
head(temp_test5)       
nfl_nn.results5<-compute(nfl_nn5,temp_test5)
nfl_nn.results5$net.result
prob5<-nfl_nn.results5$net.result
ann5<-ifelse(prob5>=.5, "run", "pass")
ann5<-as.vector(ann5)
resultsann5<-data.frame(Actual = testing5$play_type, Prediction = ann5)
tb_ann5<-table(resultsann5$Actual, resultsann5$Prediction)


confusionMatrix(tb_ann5)

tn<-subset(results_svm, Actual == "run", Prediction == "run")
tp<-subset(results_svm, Actual == "pass", Prediction == "pass")
fp<-subset(results_svm, Actual == "pass", Prediction == "run")
fn<-subset(results_svm, Actual == "run", Prediction == "pass")
ann5_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
ann_tp5<-ifelse(resultsann5$Actual == resultsann5$Prediction, 1,0)
ann_tp5<-as.data.frame(ann_tp5)
ann_tp5$model<-"ANN"
names(ann_tp5)[1] = "accuracy"
```

```{r}
#RandomForest model Down 3 Yard 0-50
random_nfl5<-randomForest(play_type ~ yards_gained + yardline_100 + down  + qtr + ydsnet ,data = training5)
random_nfl5

plot(random_nfl5)
#Prediction
ranForest_pred5<-predict(random_nfl5, testing5)

ran_tb5<-table(testing5$play_type, ranForest_pred5, dnn = c("Actual", "Prediction"))
results_rf5<-data.frame(Actual = testing5$play_type, Prediction = ranForest_pred5)

confusionMatrix(ran_tb5)


tn<-subset(results_svm, Actual == "run", Prediction == "run")
tp<-subset(results_svm, Actual == "pass", Prediction == "pass")
fp<-subset(results_svm, Actual == "pass", Prediction == "run")
fn<-subset(results_svm, Actual == "run", Prediction == "pass")
rf5_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
rf_tp5<-ifelse(results_rf5$Actual == results_rf5$Prediction, 1,0)
rf_tp5<-as.data.frame(rf_tp5)
rf_tp5$model<-"RF5"
names(rf_tp5)[1] = "accuracy"
```

```{r}
#Comparing Accuracies
compare5<-data.frame(method = c("Support Vector Machine", "Neural Network", "Random Forest"), accuracy = c(svm5_accuracy, ann5_accuracy, rf5_accuracy))
ggplot(data = compare5, mapping = aes(x= method, y = accuracy, fill = method)) +geom_bar(stat = "identity")+geom_text(aes(label = round(accuracy,4)), position = position_stack(vjust = .5), color = "white") + labs (x = "Method", y = "Accuracy", title = "Accuracy for Down 2 Yard Line 50-0")
```


# 4th Down model and prediction
#Should teams Execute a play on 4th down if they are on their own 40 with >3 yards to make a 1st down

```{r}
#subset data based on play being 4th down, yards to go less than 3 and play type. Should a run or pass play be excuted based on being between the teams own 40 and the opposing teams 40
Down4p<-subset(nflDataa, down == "4" & ydstogo < 3 & play_type == "pass" &  yardline_100 <60)

Down4r<-subset(nflDataa, down == "4" & ydstogo < 3 & play_type == "run" & yardline_100 <60)

#bind 2 data frames of run and pass 
#Logistic regression
set.seed(123)
fourth<-rbind(Down4p,Down4r)
sample1<-sample.split(fourth$yards_gained, SplitRatio = .75)
training6<-subset(fourth, sample1 == TRUE)
testing6<-subset(fourth, sample1 == FALSE)

#Logistic Model
nfl.fit<-lm(yards_gained ~ ydstogo  + yardline_100 + qtr + ydsnet + play_type, data = training1 )
summary(nfl.fit)
plot(nfl.fit)


lm_pred<-predict(nfl.fit, testing1)

results_lm1<-data.frame(Actual=testing1$yards_gained,Prediction = lm_pred)
results_lm1

average<-mean(results_lm1$Prediction)
average
#3.305929

ggplot(data = testing1, mapping = aes(x = yardline_100, y = yards_gained)) + geom_point() +geom_smooth(method = "lm", se = FALSE, color = "red", alpha = 1) + labs(x= 'Yard Line on the Field', y = 'Yards Gained')
```



```{r}
#Logistic model

#3rd and long best play type prediction 
Down3p<-subset(nflDataa, down == "3" & ydstogo < 5 & play_type == "pass" &  yardline_100 <60)

Down3r<-subset(nflDataa, down == "3" & ydstogo < 5 & play_type == "run" & yardline_100 <60)



third<-rbind(Down3p,Down3r)
#third$play_type<-as.factor(third$play_type)
third$yardline_100<-as.integer(third$yardline_100)
third$yards_gained<-as.integer(third$yards_gained)
third$ydsnet<-as.integer(third$ydsnet)
third$qtr<-as.integer(third$qtr)

third$play_type<-ifelse(third$play_type == "run", 0,1)

print(third)

set.seed(123)
sample1<-sample.split(third$play_type, SplitRatio = .75)
training7<-subset(third, sample1 == TRUE)
testing7<-subset(third, sample1 == FALSE)


glm_model<-glm(play_type~ yards_gained + yardline_100   + qtr + ydsnet, data = training7)
summary(glm_model)
plot(glm_model)


#Prediction
play_pred2<-predict(glm_model, newdata  = testing7)

glm.fit<-ifelse(play_pred2 >=0.5,1 , 0)

tb2<-table(testing7$play_type, glm.fit, dnn = c("Actual", "Prediction"))

results_glm2<-data.frame(Actual = testing7$play_type, Prediction = glm.fit)

confusionMatrix(tb2)


tn<-subset(results_glm2, Actual == "run", Prediction == "run")
tp<-subset(results_glm2, Actual == "pass", Prediction == "pass")
fp<-subset(results_glm2, Actual == "pass", Prediction == "run")
fn<-subset(results_glm2, Actual == "run", Prediction == "pass")
glm_accuracy<-(nrow(tp)+nrow(tn))/(nrow(tp)+nrow(tn)+nrow(fp)+nrow(fn))
glm_tp<-ifelse(results_glm2$Actual == results_glm2$Prediction, 1,0)
glm_tp<-as.data.frame(glm_tp)
glm_tp$model<-"RF5"
names(glm_tp)[1] = "accuracy"
```  



